---
title: "In class practice"
author: "Lowson Li"
output:
  html_notebook:
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: simplex
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: inline
---

```{r message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```
# Future value 
Formula
$$
FV = P(1+\frac{r}{m})^{nm}
$$

Example
A company should pay $20 million after 4 years. It can invest money at annual rate of 7% semiannually. How much shoud the company invest today?
$$
20,000,000 = P(1+\frac{7\%}{2})^{2\cdot4}
$$
Answer 
```{r}
20000000/((1+(0.07/2))^(1/8))
```

# Continuous compounding
Since $$\lim_{x\rightarrow \infty } (1+\frac{a}{n})^n = e^a$$
we have $$\lim_{x\rightarrow \infty } (1+\frac{r}{m})^n = e^r$$
Thus in continuous compounding, the future value should be 
$$ FV = Pe^{rn}$$
# Conversion between compunding method 
$r_1$: Annual rate with continuous compunding method
$r_2$: Equivelant rate compounded $m$ times per annum
Since 
$$
FV = P(1+\frac{r_2}{m})^{mn}=Pe^{r_1 \cdot n}
$$
We have 
$$
\begin{equation}
\begin{aligned}
r_1 &=m\ln(1+\frac{r_2}{m}) \\
r_2 &= m(e^{\frac{r_1}{m}}-1)
\end{aligned}
\end{equation}
$$
Example 
Consider interest rate of 10% with quarterly compounding. What's the equivelent rate with continunous rate?
```{r}
4*log(1+(0.1/4))
```
# Annuity 
* Ordinary annuity: **pays out** the same amount of money at the **end of month**.
* Annuity due: **receives** the same amount of money at the **begining of the month**.

## Ordinary annuity 
$$
\begin{equation}
\begin{aligned}
FV &= C +C(1+r)+...+C(1+r)^{n-1} \\
&= \sum_{i=0}^{n-1}C(1+r)^{i} \\
&= C \frac{(1+r)^n-1}{r}
\end{aligned}
\end{equation}
$$


Allowing compounded $m$ times:

$$
FV =C \frac{(1+\frac{r}{m})^{nm}-1}{\frac{r}{m}}
$$


## Annuity due 
$$
\begin{equation}
\begin{aligned}
FV &= C(1+r)+...+C(1+r)^n \\
&= \sum_{i=1}^{n}C(1+r)^{i} \\
&=C(1+r)\frac{(1+r)^n-1}{r}
\end{aligned}
\end{equation}
$$

Allowing compounded $m$ times:

$$
FV =C (1+\frac{r}{m})\frac{(1+\frac{r}{m})^{nm}-1}{\frac{r}{m}}
$$
## Present value of the general annuity
$$
P=\sum_{i=1}^{mn}C(1+r)^{-i}  = C \frac{1-(1+\frac{r}{m}^{-nm})}{\frac{r}{m}}
$$
Example
The $P$ of an annuity of $100 per annum for 5 yeats at annual rate of 6.25% is 

```{r}
# define annuity function
cal_annuity <- function(c, r, p) {
  a <- 0
  for (i in 1:p) {
    temp <- (1 + r)^(-i)
    a <- a + temp
  }
  return(c * a)
}
```
```{r}
# calculate annuity
cal_annuity(100,0.0625,5)
```
# Mortgage
Example 
Consider a 15-year, $250,000 loan at 8% interest rate, what's the monthly payment?
(Hint: use general annuity formula)
```{r}
250000/((1-((1+(0.08/12))^(-12*15)))/(0.08/12))
```

# Yields
The return of the investment.
* Bond Equivalent Yield, BEY
* Mortgage Equivalent Yield, MEY

# Zero-Coupon Bonds 
The $price$ of a zero coupon bonds that pays $F$ dollars in $n$ periods follows:
$$
Price = \frac{F}{(1+r)^n}
$$
Example: 
The interest rate is compounded semiannually. A zero coupon bond that pays the par value 20 years from now will be priced at how much of its par value?
$r=8\%,m=2,n=20$ 
$$
\frac{Price}{F} = \frac{1}{(1+\frac{r}{m})^{nm}}= \frac{1}{(1+\frac{8\%}{2})^{20\cdot2}}
$$
```{r}
1/((1+(0.08/2))^(20*2))
```
If the bond matures in 10 years instead of 20 years, the price will be 
```{r}
1/((1+(0.08/2))^(10*2))
```
# Level coupon bond

* Invester will received coupon each period.
* Denote par value $F$, annual counpon rate $c$ and the coupon received in each period $C$.
* The total number of periods is $n$.
* Assume the interest compound $m$ times annualy, the general price formula of the level coupon bond is:

$$
\begin{equation}
\begin{aligned}
P &=\frac{C}{(1+\frac{r}{m})}+...+\frac{C+F}{(1+\frac{r}{m})^n} \\
&= \frac{F}{(1+\frac{r}{m})^n}+\sum_{i=1}^{n}C(1+\frac{r}{m})^{-i} \\
&= \frac{F}{(1+\frac{r}{m})^n} +C\frac{1-(1+\frac{r}{m})^{-n}}{r}
\end{aligned}
\end{equation}
$$
where the coupon received at each period $C$ is given by $\frac{F \times c}{m} $.

Example 
15% BEY, 10 years bond with compound rate of 10% paid semiannually should sell for 
(Hint: assume the par value is 100)

* Without loss of generality, let par value $F=100$.
* Interest rate $15\%$, $m=2$
* Total number of periods $n=2 \times 10 = 20$
* Coupon received at each period $C = \frac{100 \times 0.1}{2}$.

```{r echo=TRUE}
(100*0.1/2)*((1-(1+(0.15/2))^(-10*2))/(0.15/2))+(100/(1+(0.15/2))^(10*2))
```

Example 
15 years, 9% coupon rate 
```{r}
# define yield-price function 
cal_price <- function(n, c, m) {
  p <- c()
  s <- seq(0.075, 0.105, 0.005)
  for (i in 1:length(s)) {
    p[i] <- 100 / ((1 + s[i] / m)^(n * m)) + (100 * c / m) * ((1 - (1 + (s[i] / m))^(-n * m)) / (s[i] / m))
  }
  return(p)
}
```
```{r}
# calculate the prices under different yield 
cal_price(15, 0.09, 2)
```

# Bond price when sell
## Accrued interest 
* Let $\omega$ be the remaining portions of coupon when sell.
The accrued interest (應計利息) will be 

$$
AI = C \times (1-\omega)
$$
### Method to calculate the days (TBD)
## Present value when sell
* In UK and US, the quoted price does not inclue the accrued interest, this is call clean price / flat price. 

The present value of the bonds when sell:

$$
\begin{equation}
\begin{aligned}
PV &=clean \ price + AI = \frac{C}{(1+\frac{r}{m})^{\omega}}+...+\frac{C+F}{(1+\frac{r}{m})^{\omega+n-1}} \\
&= \frac{F}{(1+\frac{r}{m})^{\omega+n-1}}+\sum_{i=0}^{n-1} \frac{C}{(1+\frac{r}{m})^{\omega+i}} \\
\end{aligned}
\end{equation}
$$
Example(TBD):

# Bond price volatility
## Duration 
### Macaulay duration (MD)

(insert formula)

### Modified duration 

(insert formula)

Example: Modified duration 11.54, change in yield 0.1%, the price should **decrease 1.154%**

### Effective duration 

(insert formula)

## Convexity 

(insert formula)

## Immutization

# Spot rate 
Compare the different bonds to the present base.
Change the coumpound rate in coupon bonds to sigle rate. 
Useful when you have investment profolio.

(insert formula)

# Forward rate 

(insert formula)
(insert example)

# Options 
## Definition 

## Types of options

## Payoff 
### Convinient convetions

* C: call value (買權權利金)
* P:put value (賣權權利金)
* X/K: strike price (履約價)
* S: stock price (標的物現貨價格)
* D: dividends (是否有發股利)

### Mathematic representation 
* The payoff of a call at expriation is 

$$
C = \max (0,S-X)
$$

When the **stock price is lower than the strike pirce** (S<X), the investor will give up call option.

* The payoff of a pu at expriation is 

$$
P = \max (0,X-S)
$$

When the **stock price is higher than the strike pirce** (S>X), the investor will give up put option.

## Strategy 

(insert image)

## Terminology

|Option|Strike Price|Stock Price| Terminology|
|:----:|:----:|:----:|:----:|
|Call|45|40|**Out**-of-the-money(放棄)|
|Call|40|40|At-of-the-money|
|Call|35|40|**In**-of-the-money(賺錢)|
|Put|80|100|**Out**-of-the-money(放棄)|
|Put|100|100|At-of-the-money|
|Put|130|100|**In**-of-the-money(賺錢)|

## Straddle vs. strangle
(strategy)

## Option pricing model 

## Black-Scholes Model
Also called Black-Scholes-Merton Model

### Basic assumptions

|Assumptions|Violation|
|:----:|:----:|
|資本市場完美、沒有稅或交易成本、股票價格沒有上下限、任何五票可以無限分割、可無限賣空|景氣不好的時候政府可能會禁止賣空|
|股價取對數後符合常態分配|有可能會集中在某些區間|
|若令股價過程為常態分配，股價可能值將從負無窮大到正無窮大|真實股價沒有負值|
|無風險利率、股票報酬率的波動度為常數|FED常更改無風險利率通常隨季度調整、波動不為常數|
|選擇權存續期間，股票不發放任何股利||
|歐式選擇權，只能在到期日時履約|還有美式選擇權|
|標的股票不會違約|有可能股票會下市|
|股價是連續的，不會有突發性股價跳空的情況發生|最近常跳空|

### Formula 


$$
\begin{equation}
\begin{aligned}
c &= S_{0}N(d_{1})-Ke^{-rT}N(d_{2}) \\
p &= Ke^{-rT}N(-d_{2})-S_{0}N(-d_{1}) \\

where \ \  d_{1} &=  \frac{\ln(S_{0}/K)+(r+\sigma^{2}/2)T}{\sigma \sqrt(T)} \\
d_{2} &=  \frac{\ln(S_{0}/K)+(r-\sigma^{2}/2)T}{\sigma \sqrt(T)} = \sigma \sqrt(T) \\
\end{aligned}
\end{equation}
$$


* Concept: 現價($S_{0}$) 服從常態分佈，履約價($K$)要折現回來，服從常態分佈。$\sigma$為標的物波動率(估計)。$r$為無風險利率。


### Put Call Parity 

* 買賣權平價關係

$$
\begin{equation}
\begin{aligned}
C+Xe^{-rT} &= P+S \\
where \ \  C &= call \ premium \\
Xe^{-rT} &= present \ value \ of \ the \ strike \\
P &= put \ premium \\ 
S &= the \ current \ price \ of \ the \ underline \\ 
\end{aligned}
\end{equation}
$$
* Geometric Bronian Motion 

$$
\mathrm{d}S_{t} := \mu S_{t}\mathrm{d}d_{t} +\sigma S_{t} \mathrm{d}B_{t}, 0\leq t \leq T
$$
where $S_{t}$ is the stock price at time $t$,  $\mu$ is the drift rate(股價年收益率的期望值),$\sigma$ is the volatility of the price in years, $B_{t}$ is the Brownian motion. Notice that $\mu, \sigma$ are constants.

## Option greek 

|Greek|Measure|Affected by|Affects|
|:----|:----|:----|:----|
|Delta|Impact of underlying price move(per $1)|Time until expiration, volatility, gamma, price|Option Price|
|Gamma|Impact of underlying price move(per $1)|Time until expiration, volatility, price|Delta|
|Theta|Impact of passage of time(per day)|Time until expiration, price, volatility|Option Price|
|Vega|Impact of volatility change(per 1%)|Time until expiration, volatility|Option Price|
|Rho|Impact of interest rate change (per 1%)|Interest rate, time until expiration|Option Price|
