---
title: "In class practice"
author: "Lowson Li"
output:
  html_notebook:
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: simplex
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: inline
---

```{r message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```
# Future value 
Formula
$$
FV = P(1+\frac{r}{m})^{nm}
$$

Example
A company should pay $20 million after 4 years. It can invest money at annual rate of 7% semiannually. How much shoud the company invest today?
$$
20,000,000 = P(1+\frac{7\%}{2})^{2\cdot4}
$$
Answer 
```{r}
20000000/((1+(0.07/2))^(1/8))
```

# Continuous compounding
Since $$\lim_{x\rightarrow \infty } (1+\frac{a}{n})^n = e^a$$
we have $$\lim_{x\rightarrow \infty } (1+\frac{r}{m})^n = e^r$$
Thus in continuous compounding, the future value should be 
$$ FV = Pe^{rn}$$
# Conversion between compunding method 
$r_1$: Annual rate with continuous compunding method
$r_2$: Equivelant rate compounded $m$ times per annum
Since 
$$
FV = P(1+\frac{r_2}{m})^{mn}=Pe^{r_1 \cdot n}
$$
We have 
$$
\begin{equation}
\begin{aligned}
r_1 &=m\ln(1+\frac{r_2}{m}) \\
r_2 &= m(e^{\frac{r_1}{m}}-1)
\end{aligned}
\end{equation}
$$
Example 
Consider interest rate of 10% with quarterly compounding. What's the equivelent rate with continunous rate?
```{r}
4*log(1+(0.1/4))
```
# Annuity 
* Ordinary annuity: **pays out** the same amount of money at the **end of month**.
* Annuity due: **receives** the same amount of money at the **begining of the month**.

## Ordinary annuity 
$$
\begin{equation}
\begin{aligned}
FV &= C +C(1+r)+...+C(1+r)^{n-1} \\
&= \sum_{i=0}^{n-1}C(1+r)^{i} \\
&= C \frac{(1+r)^n-1}{r}
\end{aligned}
\end{equation}
$$


Allowing compounded $m$ times:

$$
FV =C \frac{(1+\frac{r}{m})^{nm}-1}{\frac{r}{m}}
$$


## Annuity due 
$$
\begin{equation}
\begin{aligned}
FV &= C(1+r)+...+C(1+r)^n \\
&= \sum_{i=1}^{n}C(1+r)^{i} \\
&=C(1+r)\frac{(1+r)^n-1}{r}
\end{aligned}
\end{equation}
$$

Allowing compounded $m$ times:

$$
FV =C (1+\frac{r}{m})\frac{(1+\frac{r}{m})^{nm}-1}{\frac{r}{m}}
$$
## Present value of the general annuity
$$
P=\sum_{i=1}^{mn}C(1+r)^{-i}  = C \frac{1-(1+\frac{r}{m}^{-nm})}{\frac{r}{m}}
$$
Example
The $P$ of an annuity of $100 per annum for 5 yeats at annual rate of 6.25% is 

```{r}
# define annuity function
cal_annuity <- function(c, r, p) {
  a <- 0
  for (i in 1:p) {
    temp <- (1 + r)^(-i)
    a <- a + temp
  }
  return(c * a)
}
```
```{r}
# calculate annuity
cal_annuity(100,0.0625,5)
```
# Mortgage
Example 
Consider a 15-year, $250,000 loan at 8% interest rate, what's the monthly payment?
(Hint: use general annuity formula)
```{r}
250000/((1-((1+(0.08/12))^(-12*15)))/(0.08/12))
```

# Yields
The return of the investment.
* Bond Equivalent Yield, BEY
* Mortgage Equivalent Yield, MEY

# Zero-Coupon Bonds 
The $price$ of a zero coupon bonds that pays $F$ dollars in $n$ periods follows:
$$
Price = \frac{F}{(1+r)^n}
$$
Example: 
The interest rate is compounded semiannually. A zero coupon bond that pays the par value 20 years from now will be priced at how much of its par value?
$r=8\%,m=2,n=20$ 
$$
\frac{Price}{F} = \frac{1}{(1+\frac{r}{m})^{nm}}= \frac{1}{(1+\frac{8\%}{2})^{20\cdot2}}
$$
```{r}
1/((1+(0.08/2))^(20*2))
```
If the bond matures in 10 years instead of 20 years, the price will be 
```{r}
1/((1+(0.08/2))^(10*2))
```
# Level coupon bond
$$
\begin{equation}
\begin{aligned}
P &=\frac{C}{(1+r)}+...+\frac{C+F}{(1+r)^n} \\
&= \frac{F}{(1+r)^n}+\sum_{i=1}^{n}C(1+r)^{-i} \\
&= \frac{F}{(1+r)^n} +C\frac{1-(1+r)^{-n}}{r}
\end{aligned}
\end{equation}
$$
General case with $m$ coupon per annum. 
$$
\begin{equation}
\begin{aligned}
P &= \frac{F}{(1+\frac{r}{m})^{nm}} +C\frac{1-(1+\frac{r}{m})^{-nm}}{\frac{r}{m}} 
\end{aligned}
\end{equation}
$$
Define coupon $C$  is given by the par value $F$, the coupon rate $c$ and the compounding frequency $m$, such that 
$$
C=\frac{Fc}{m}
$$

Example 
15% BEY, 10 years bond with compound rate of 10% paid semiannually should sell for 
(Hint: assume the par value is 100)
```{r}
(100*0.1/2)*((1-(1+(0.15/2))^(-10*2))/(0.15/2))+(100/(1+(0.15/2))^(10*2))
```

Example 
15 years, 9% coupon rate 
```{r}
# define yield-price function 
cal_price <- function(n, c, m) {
  p <- c()
  s <- seq(0.075, 0.105, 0.005)
  for (i in 1:length(s)) {
    p[i] <- 100 / ((1 + s[i] / m)^(n * m)) + (100 * c / m) * ((1 - (1 + (s[i] / m))^(-n * m)) / (s[i] / m))
  }
  return(p)
}
```
```{r}
# calculate the prices under different yield 
cal_price(15, 0.09, 2)
```

